---
output: github_document
---

# MRchecks

## Overview

`MRchecks` is a package that performs all the analyses in a forthcoming paper on how time-varying exposures can bias Mendelian randomization (MR) analyses. With the functions in this package you can graph and estimate how the association of a genetic variant with a phenotype changes with age. If you're concerned that the association changes enough that it may bias your MR estimate, there is also a function that will estimate the potential bias under a number of exposure windows.

## Installation

Installation requires devtools which can be installed with `install.packages("devtools")`. Once you have `devtools` installed you can install MRchecks with:

``` r
devtools::install_github("jalabrecque/MRchecks")
```

## Usage

The package comes with a small test data set called `test_data`:

```{r}
library(MRchecks)
data(test_data)
```

The data set contains the minimal variables required: 

```{r}
head(test_data)
```

An age variable (age), a factor variable with the genetic data (SNP) a phenotype that will be used as the exposure in the MR analysis (y) and here, optionally, a covariate is included  (c).

### Run model

We can first use the `SNPxAGE_model` function to fit the model we will use as the input for the other functions: 
```{r}
SNPxAGE_model_output <- SNPxAGE_model(data = test_data,
                                      SNP = "SNP",
                                      phenotype = "y",
                                      age = "age",
                                      k = 3,
                                      covars="c")
```

The argument k specifies how many internal knots should be used.

### Plotting age-varying genetic associations

We can plot the relationship between age and the phenotype by genetic variant as well as a plot of additive genetic effects by age:

```{r}
SNPxAGE_plot(SNPxAGE_model_output)
```


### Plasmode simulation for potential bias

Looking at the output of this plot we might be concerned that the genetic association with the phenotype varies enough that bias is a real concern. Using `SNPxAGE_bias` we can check what the bias would be if all the MR assumptions hold:

```{r}
SNPxAGE_bias(SNPxAGE_model_output,rep = 10,age_set = 66)
```

